Postural developments modulate children's visual access to social information

### 2018-02-01 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
mikes last edis

author-information: > 
    \author{{\large \bf Alessandro Sanchez\thanks{These authors contributed equally to this work.}} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Bria Long\footnotemark[1]}  \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Allison M. Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

@@

---
\newcommand{\wrapmf}[1]{#1}

@@

Recent technological developments allow for testing of this hypothesis by documenting the experiences of infants and children from their own perspective. By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children -- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008]. Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2013statistical]. This perspective changes over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016]. These changes may be driven by – or independent from – the concurrent motoric changes described above.

@@

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation. This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and a NSF #1714726 to BLL. An earlier version of this work was presented to the Cognitive Science Society in @frank2013.

>>>>

author-information: > 
    \renewcommand*{\thefootnote}{\fnsymbol{footnote}} 
    \author{{\large \bf Alessandro Sanchez\thanks{These authors contributed equally to this work.}} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Bria Long\footnotemark[1]}  \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Allison M. Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

@@

---
\newcommand{\wrapmf}[1]{#1}
\renewcommand*{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

@@

Recent technological developments allow for testing of this hypothesis by documenting the experiences of infants and children from their own perspective. By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children -- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008]. Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggests that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2013statistical]. This perspective changes over the first two years of life, as visual input transitions from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016]. It is uncertain how much these changes are driven by the concurrent motoric changes described above versus by other developmental factors.

@@

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation. This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and NSF #1714726 to BLL. An earlier version of this work was presented to the Cognitive Science Society in @frank2013.

### 2018-02-01 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
mikes last edis

@@

\newcommand{\wrapmf}[1]{#1}

@@

These motor improvements lead to developmental cascades that impact
children's emerging social, cognitive, and linguistic abilities in
various ways. For example, postural changes impact how children interact
with their mothers: walking (vs.~crawling) infants make different kinds
of object-related bids for attention from their mothers and tend to hear
more action directed statements (e.g., ``open it'') (Karasik,
Tamis-LeMonda, \& Adolph, 2014). Further, in an observational study,
Walle \& Campos (2014) found that children who were able to walk had
both higher receptive and productive vocabularies. Thus, children's
ability independently locomote may change their ability to access social
information (e.g., faces, gaze) and in turn accelerate their learning.

Recent technological developments allow for testing of this hypothesis
by documenting the experiences of infants and children from their own
perspective. By using head-mounted cameras, researchers have begun to
record the visual experiences of infants and children -- which even for
walking children are extremely different from the adult perspective (and
not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; Franchak et al., 2011; Yoshida \& Smith, 2008).
Children's views tend to be more restricted and dominated by objects and
hands (Yoshida \& Smith, 2008), and computational and empirical work
suggest that this restricted viewpoint may be more effective for
learning objects and their labels than the comparable adult perspective
(Bambach, Crandall, Smith, \& Yu, 2017; Yurovsky, Smith, \& Yu, 2013).
This perspective changes over the first two years of life, as views
transition from primarily containing close up views of faces to
capturing views of hands paired with the objects they are acting on
(Fausey, Jayaraman, \& Smith, 2016). These changes may be driven by --
or independent from -- the concurrent motoric changes described above.

Here, we directly examine whether postural and locomotor developments
change the availability of social information -- the presence of faces
and hands. To do so, we recorded the visual experience of a group of
infants in three age ranges (8,12, and 16 months) using head-mounted
cameras during a brief laboratory free-play session; children's posture
and orientation relative to their caregiver were also recorded from a
third-person perspective and hand-annotated. We then capitalize on
recent improvements in face and pose detection algorithms (Cao, Simon,
Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
frequencies of faces and hands (using wrists as a proxy for the latter)
in the child's visual environment, both overall and relative to naming
events by their caregivers. We hypothesized that there would be
differential access to social information based on children's postural
developments: crawling infants would see fewer faces because they would
primarily be looking at the ground, while walking toddlers would have
access to a richer visual landscape, thus providing children with
greater access to the social information in their environment.

@@

We used a small, head-mounted camera (``headcam'') that was constructed
from a MD80 model camera attached to a soft elastic headband. Videos
captured by the headcam were 720x480 pixels with 25 frames per
second.\footnote{Detailed instructions for creating this headcam can be
  found at
  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
A fisheye lens was attached to the camera to increase the view angle
from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
\ref{fig:headcam}, bottom).

@@

To evaluate face detector performance, we hand-labeled a ``gold set'' of
labeled frames. To account for the relatively rare appearance of faces
in the dataset, we hand-labeled two types of samples: a sample
containing a high density of faces (half reported by MTCNN, half by
OpenPose) and a random sample from the remaining frames. Each sample was
comprised of an equal number of frames taken from each child's video.
For wrist detections, the ``gold set'' was constructed in the same
manner, except frames with a high density of wrists came only from
detections made by OpenPose. Faces were classified as present if at
least half of the face was showing; wrists were classified as present if
any part of the wrist was showing. Precision (hits / hits + false
alarms), recall (hits / hits + misses), and F-score (harmonic mean of
precision and recall) were calculated for all detectors and are reported
in Table 2.

For face detection, MTCNN outperformed OpenPose when taking into account
only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose). Although
MTCNN and OpenPose performed comparably with the random sample, MTCNN
performed better on the high density sample (specifically looking at
precision), suggesting that OpenPose generated more false positives than
MTCNN. ViolaJones performed quite poorly relative to the other
detectors, especially with respect to the random sample. We thus use
MTCNN detections in the following analyses. For wrist detection,
OpenPose performed moderately well (F = 0.74) with relatively high
precision but low recall on the randomly sampled frames (see Table 2).
We thus analyze wrist detections, with the caveat that we are likely
underestimating the proportion of hands in the dataset.

@@

We examined the proportion of face and hand detections across age (see
Figure \ref{fig:detByAge}). We observed a slight U-shaped function in
face detections, such that 12-month-olds appeared to have visual access
to slightly fewer faces faces than 8 or 16-month-olds; conversely, hand
detections perhaps appeared to generally increase with age.

@@

To formalize these observations, we fit two generalized linear
mixed-effect models for the proportion of faces and the proportion of
hands infants saw in each posture and orientation, with participant's
age, orientation, and posture as fixed effects and with random slopes
for infant's orientation. Random slopes for posture or interactions
between our fixed effects caused the models to fail to converge.

@@

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
recruitment, data collection, and annotation. This work was funded by a
Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
and a NSF \#1714726 to BLL. An earlier version of this work was
presented to the Cognitive Science Society in Frank et al. (2013).

>>>>

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

@@

\newcommand{\wrapmf}[1]{#1}
\renewcommand*{\thefootnote}{\arabic{footnote}}

\setcounter{footnote}{0}

@@

These motor improvements lead to developmental cascades that impact
children's emerging social, cognitive, and linguistic abilities in
various ways. For example, postural changes impact how children interact
with their mothers: walking (vs.~crawling) infants make different kinds
of object-related bids for attention from their mothers and tend to hear
more action directed statements (e.g., ``open it'') (Karasik,
Tamis-LeMonda, \& Adolph, 2014). Further, in an observational study,
Walle \& Campos (2014) found that children who were able to walk had
both higher receptive and productive vocabularies. Thus, children's
ability to independently locomote may change their ability to access
social information (e.g., faces, gaze) and in turn accelerate their
learning.

Recent technological developments allow for testing of this hypothesis
by documenting the experiences of infants and children from their own
perspective. By using head-mounted cameras, researchers have begun to
record the visual experiences of infants and children -- which even for
walking children are extremely different from the adult perspective (and
not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; Franchak et al., 2011; Yoshida \& Smith, 2008).
Children's views tend to be more restricted and dominated by objects and
hands (Yoshida \& Smith, 2008), and computational and empirical work
suggests that this restricted viewpoint may be more effective for
learning objects and their labels than the comparable adult perspective
(Bambach, Crandall, Smith, \& Yu, 2017; Yurovsky, Smith, \& Yu, 2013).
This perspective changes over the first two years of life, as visual
input transitions from primarily containing close up views of faces to
capturing views of hands paired with the objects they are acting on
(Fausey, Jayaraman, \& Smith, 2016). It is uncertain how much these
changes are driven by the concurrent motoric changes described above
versus by other developmental factors.

Here, we directly examine whether postural and locomotor developments
change the availability of social information -- the presence of faces
and hands. To do so, we recorded the visual experience of a group of
infants in three age ranges (8, 12, and 16 months) using head-mounted
cameras during a brief laboratory free-play session; children's posture
and orientation relative to their caregiver were also recorded from a
third-person perspective and hand-annotated. We then capitalize on
recent improvements in face and pose detection algorithms (Cao, Simon,
Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
frequencies of faces and hands (using wrists as a proxy for the latter)
in the child's visual environment, both overall and relative to naming
events by their caregivers. We hypothesized that there would be
differential access to social information based on children's postural
developments: crawling infants would see fewer faces because they would
primarily be looking at the ground, while walking toddlers would have
access to a richer visual landscape, thus providing children with
greater access to the social information in their environment.

@@

We used a small, head-mounted camera (``headcam'') that was constructed
from a MD80 model camera attached to a soft elastic headband. Videos
captured by the headcam were 720x480 pixels with 25 frames per
second.\footnote{Detailed instructions for creating this headcam can be
  found at
  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
A fisheye lens was attached to the camera to increase the view angle
from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
\ref{fig:headcam}, above).

@@

To evaluate face detector performance, we hand-labeled a ``gold set'' of
labeled frames. To account for the relatively rare appearance of faces
in the dataset, we hand-labeled two types of samples: a sample
containing a high density of faces (half reported by MTCNN, half by
OpenPose) and a random sample from the remaining frames. Each sample was
comprised of an equal number of frames taken from each child's video.
For wrist detections, the ``gold set'' was constructed in the same
manner, except frames with a high density of wrists came only from
detections made by OpenPose. Faces were classified as present if at
least half of the face was showing; wrists were classified as present if
any part of the wrist was showing. Precision (hits / hits + false
alarms), recall (hits / hits + misses), and F-score (harmonic mean of
precision and recall) were calculated for all detectors and are reported
in Table 1.

For face detection, MTCNN outperformed OpenPose when taking into account
only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose). Although
MTCNN and OpenPose performed comparably with the random sample, MTCNN
performed better on the high density sample (specifically looking at
precision), suggesting that OpenPose generated more false positives than
MTCNN. ViolaJones performed quite poorly relative to the other
detectors, especially with respect to the random sample. We thus use
MTCNN detections in the following analyses. For wrist detection,
OpenPose performed moderately well (F = 0.74) with relatively high
precision but low recall on the randomly sampled frames (see Table 1).
We thus analyze wrist detections, with the caveat that we are likely
underestimating the proportion of hands in the dataset.

@@

We examined the proportion of face and hand detections across age (see
Figure \ref{fig:detByAge}). We observed a slight U-shaped function in
face detections, such that 12-month-olds appeared to have visual access
to slightly fewer faces than 8 or 16-month-olds; conversely, hand
detections perhaps appeared to generally increase with age.

@@

To formalize these observations, we fit two generalized linear
mixed-effect models for the proportion of faces and the proportion of
hands infants saw in each posture and orientation, with participant's
age, orientation, and posture as fixed effects and with random slopes
for infants' orientation. Random slopes for posture or interactions
between our fixed effects caused the models to fail to converge.

@@

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
recruitment, data collection, and annotation. This work was funded by a
Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
and NSF \#1714726 to BLL. An earlier version of this work was presented
to the Cognitive Science Society in Frank et al. (2013).
